---
title: "Child Patients with Covid-19 and Cystic Fibrosis"
author: "Alfonso Vieyra, Brandon Huett, Allison Cher, Brittnie Villasenor, Ernesto Collazo"
date: '2022-06-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(scales)
```

## EDA Analysis

```{r morbidity Tables}
#table with severity, cf and comorbidities NOTE: Changed names for readability
morb_df <- d %>%
  select(has_cf, COVIDseverity, starts_with('comorb'))


colnames(morb_df) <- c('has_cf', 'COVIDseverity', 'bronchiectasis',
                        'pulmonary_disease', 'asthma', 'nasal_polyps',
                        'hemoptysis', 'pneumothorax', 'resp_failure', 
                        'malnutrition', 'other_nutri_deficiency', 
                        'type1_diabetes', 'type2_diabetes', 'obesity', 
                        'liver_disease', 'other_GI_notLiver', 
                        'essential_hypertension', 'hypertensive_heart_disease',
                        'chronic_kidney_disease', 'nicotine_dependence', 
                        'heart_failure', 'ischemic_heart_disease', 
                        'lung_transplant')

#table of non cystic fibrosis patients
morb_noCF <- morb_df  %>%
  filter(has_cf == 0)

#table of cystic fibrosis patients
morb_CF <- morb_df %>%
  filter(has_cf == 1)
```

```{r funcy}
#created a function to calculate the proportions in a 1D vector
prop1 <- function(arr, denom) {
  count = 1
  for (x in arr)
  {
    arr[count] = round(x / denom, 2)
    count = count + 1
  }
  return(arr)
}
```

```{r Severity Levels noCF}

#Tables of severity symptoms level 0
severity0_noCF <- morb_noCF %>%
  filter(COVIDseverity == 0) %>%
  select(-has_cf, -COVIDseverity)

indx0 <- sapply(severity0_noCF, is.factor)
severity0_noCF[indx0] <- lapply(severity0_noCF[indx0], 
                           function(x) as.numeric(as.character(x)))H <- 

#Tables of severity symptoms level 1
severity1_noCF <- morb_noCF %>%
  filter(COVIDseverity == 1) %>%
  select(-has_cf, -COVIDseverity)

indx1 <- sapply(severity1_noCF, is.factor)
severity1_noCF[indx1] <- lapply(severity1_noCF[indx1], 
                           function(x) as.numeric(as.character(x)))

#Tables of severity symptoms level 2 
severity2_noCF <- morb_noCF %>%
  filter(COVIDseverity == 2) %>%
  select(-has_cf, -COVIDseverity)

indx2 <- sapply(severity2_noCF, is.factor)
severity2_noCF[indx2] <- lapply(severity2_noCF[indx2], 
                           function(x) as.numeric(as.character(x)))

#Vectorized severities of counts
asymptomatic <- colSums(severity0_noCF)
mild_symptoms <- colSums(severity1_noCF)
severe_symptoms <- colSums(severity2_noCF)

#created new df based on counts of morbidities across severity lvls 
library(reshape2)

sev_lvls_noCF <- rbind(asymptomatic, mild_symptoms, severe_symptoms)
sev_lvls_noCF[1, ] <- prop1(sev_lvls_noCF[1,], 296236)
sev_lvls_noCF[2, ] <- prop1(sev_lvls_noCF[2,], 34015)
sev_lvls_noCF[3, ] <- prop1(sev_lvls_noCF[3,], 20743)

melted_sev_lvls_noCF <- melt(sev_lvls_noCF)
colnames(melted_sev_lvls_noCF) <- c('severity_lvl', 'morbidity','proportion')
```


```{r visual non-CF}
options(scipen = 999)
ggplot(melted_sev_lvls_noCF, aes(x=morbidity, y = proportion, fill = proportion)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  labs( x = 'Morbidities', y = 'Total Reported Conditions') +
  ggtitle('Dominant Conditions in Non-CF patients with Covid 19') +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~severity_lvl) + 
  theme(axis.text.x = element_text(angle=45))
```

```{r Severity Levels CF}
#Tables of severity symptoms level 0
severity0_CF <- morb_CF %>%
  filter(COVIDseverity == 0) %>%
  select(-has_cf, -COVIDseverity)

indx0_cf <- sapply(severity0_CF, is.factor)
severity0_CF[indx0_cf] <- lapply(severity0_CF[indx0_cf], 
                           function(x) as.numeric(as.character(x)))

#Tables of severity symptoms level 1
severity1_CF <- morb_CF %>%
  filter(COVIDseverity == 1) %>%
  select(-has_cf, -COVIDseverity)

indx1_cf <- sapply(severity1_CF, is.factor)
severity1_CF[indx1_cf] <- lapply(severity1_CF[indx1_cf], 
                           function(x) as.numeric(as.character(x)))

#Tables of severity symptoms level 2 
severity2_CF <- morb_CF %>%
  filter(COVIDseverity == 2) %>%
  select(-has_cf, -COVIDseverity)

indx2_cf <- sapply(severity2_CF, is.factor)
severity2_CF[indx2_cf] <- lapply(severity2_CF[indx2_cf], 
                           function(x) as.numeric(as.character(x)))

#Vectorized severities of counts
asympt_mild <- colSums(severity0_CF)
moderate_symptoms <- colSums(severity1_CF)
severe_symptoms <- colSums(severity2_CF)

#created new df based on counts of morbidities across severity lvls 
sev_lvls_CF <- rbind(asympt_mild, moderate_symptoms, severe_symptoms)
sev_lvls_CF[1,] <- prop1(sev_lvls_CF[1,], 8454)
sev_lvls_CF[2,] <- prop1(sev_lvls_CF[2,], 1216)
sev_lvls_CF[3,] <- prop1(sev_lvls_CF[3,], 646)

melted_sev_lvls_CF <- melt(sev_lvls_CF)
colnames(melted_sev_lvls_CF) <- c('severity_lvl', 'morbidity','proportion')
```

```{r visual CF}
ggplot(melted_sev_lvls_CF, aes(x=morbidity, y = proportion, fill = proportion)) +
  geom_bar(stat = 'identity') +
  coord_flip()+
  labs( x = 'Morbidities', y = 'Proportion of Conditions') +
  ggtitle('Dominant Conditions in CF patients with Covid 19') +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~severity_lvl) + 
  theme(axis.text.x = element_text(angle=45))
```

```{r cf non cf}
has_CF <- morb_CF %>%
  select(-has_cf, -COVIDseverity)

indx_hasCF <- sapply(has_CF, is.factor)
has_CF[indx_hasCF] <- lapply(has_CF[indx_hasCF], 
                           function(x) as.numeric(as.character(x)))

has_noCF <- morb_noCF %>%
  select(-has_cf, -COVIDseverity)

indx_hasNoCF <- sapply(has_noCF, is.factor)
has_noCF[indx_hasNoCF] <- lapply(has_noCF[indx_hasNoCF], 
                           function(x) as.numeric(as.character(x)))

cyst_fibrosis_patients <- colSums(has_CF)
noncyst_fibrosis_patients <- colSums(has_noCF)

cf_lvls <- rbind(cyst_fibrosis_patients, noncyst_fibrosis_patients)
cf_lvls[1,] <- prop1(cf_lvls[1,], 10316)
cf_lvls[2,] <- prop1(cf_lvls[2,], 350994)

melted_cf_lvls <- melt(cf_lvls)
colnames(melted_cf_lvls) <- c('has_CF', 'morbidity','proportion')
```

```{r visuals cf lvls}
ggplot(melted_cf_lvls, aes(x=morbidity, y = proportion, fill = proportion)) +
  geom_bar(stat = 'identity') +
  coord_flip()+
  labs( x = 'Morbidities', y = 'Proportion of Conditions') +
  ggtitle('Dominant Conditions in CF and non-CF patients with Covid 19') +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~has_CF) + 
  theme(axis.text.x = element_text(angle=45))
```

```{r # of morbidities in patients}
#number of mordidities in each patient
num_morb_perCF <- table(rowSums(has_CF))
num_morb_perNonCF <- table(rowSums(has_noCF))

#Assign proportions for the number of comorbidities per patient
num_morb_perCF <- prop1(num_morb_perCF, 9337)
num_morb_perNonCF <- prop1(num_morb_perNonCF, 916615)
```

```{r unique patient df}
unique_patient_df <- 
  d[order(d$servicedate),] %>%
  distinct(personid, .keep_all = TRUE)
```

```{r tenant}
h <- d %>% 
  group_by(d$tenant) %>%
  filter(n()>17500)

ggplot(h) +
  geom_bar(mapping = aes(x = tenant, fill = COVIDseverity), )
```

```{r encounterType}
ggplot(data = d) +
  geom_bar(mapping = aes(x = encountertype, fill = COVIDseverity), ) + 
  coord_flip()
```

CPAP - Continuous positive airway pressure is a form of positive airway pressure 
ventilation in which a constant level of pressure greater than atmospheric 
pressure is continuously applied to the upper respiratory tract of a person.

BiPAP - bilevel positive airway pressure. It is commonly known as “BiPap” or 
“BPap.” It is a type of ventilator—a device that helps with breathing. 
During normal breathing, your lungs expand when you breathe in.

In extracorporeal membrane oxygenation (ECMO), blood is pumped outside of your 
body to a heart-lung machine that removes carbon dioxide and sends oxygen-filled 
blood back to tissues in the body.

```{r contingency table of treatment one across severity}
#
ecmo <- d %>%
  select(COVIDseverity, ECMO) %>%
  table()
ecmo


#high flow nasal cannula
hfnc <- d %>%
  select(COVIDseverity, HFNC) %>%
  table()
hfnc


#nasal cannula
nc <- d %>%
  select(COVIDseverity, NC) %>%
  table()
nc


#bilevel positive airway pressure/Continuous positive airway pressure
bipap_cpap <- d %>%
  select(COVIDseverity, bipapCpapNonInvasive) %>%
  table()
bipap_cpap


mech_vent <- d %>%
  select(COVIDseverity, invasiveMechVentilation) %>%
  table()
mech_vent
```
NOTE: Every treatment therapy was associated with either level one severty or 
level two severity or both.
