---
title: "Child Patients with Covid-19 and Cystic Fibrosis"
author: "Alfonso Vieyra, Brandon Huett, Allison Cher, Brittnie Villasenor, Ernesto Collazo"
date: '2022-06-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(scales)

# Need to always load the package though
library(epitools)
library(rmeta)
library(pROC)
library(nnet)
```

We first create a sub data frame consisting only of patients who reported 
their first visit and columns consisting of the treatments, has_cf and 
COVIDseverity.

```{r DataFrame}
unique_patient_df <- 
  d[order(d$servicedate),] %>%
  distinct(personid, .keep_all = TRUE)

treatments_df <- unique_patient_df %>%
  select(has_cf, COVIDseverity, 37:41)

```

## contingency tables of treatments as X and has_cf as Y

```{r ECMO}
ecmo <- unique_patient_df %>%
  select(has_cf, ECMO) %>%
  table()
summary(ecmo)
chisq.test(ecmo)
```

Here we observe whether or not having cystic fibrosis is dependent on having 
ECMO as a treatment. The p value = 1 and thus we can conclude that patients on 
ECMO is not indicative of whether or not they have cystic fibrosis.

```{r HFNC}
#high flow nasal cannula
hfnc <- unique_patient_df %>%
  select(has_cf, HFNC) %>%
  table()
summary(hfnc)
chisq.test(hfnc)
```

Here we observe whether or not having cystic fibrosis is dependent on having 
High Nasal Flow Canula as a treatment. The p value is approximately zero and 
thus we can conclude that patients on HFNC is significant in determining 
whether or not a patient has cystic fibrosis.

```{r NC}
#nasal cannula
nc <- unique_patient_df %>%
  select(has_cf, NC) %>%
  table()
summary(nc)
chisq.test(nc)
```

Here we observe whether or not having cystic fibrosis is dependent on having 
Nasal Canula as a treatment. The p value is approximately zero and thus we can 
conclude that patients on NC is significant in determining whether or not a 
patient has cystic fibrosis.

```{r BiPap and CPap}
#bilevel positive airway pressure/Continuous positive airway pressure
bipap_cpap <- unique_patient_df %>%
  select(has_cf, bipapCpapNonInvasive) %>%
  table()
summary(bipap_cpap)
chisq.test(bipap_cpap)
```

Here we observe whether or not having cystic fibrosis is dependent on having 
bilevel positive airway pressure/Continuous positive airway pressure as a 
treatment. The p value = 0.016 and thus we can conclude that patients on 
bipap_cpap is significant in determining whether or not a patient has cystic 
fibrosis.

```{r Invasive Mech}
mech_vent <- unique_patient_df %>%
  select(has_cf, invasiveMechVentilation) %>%
  table()
summary(mech_vent)
chisq.test(mech_vent)
```

Here we observe whether or not having cystic fibrosis is dependent on having 
invasive mechanical procedure as a treatment. The p value = 0.006 and thus we 
can conclude that patients on mech_vent is significant in determining whether or 
not a patient has cystic fibrosis.

## Ordinal multinomial logistic Model of treatments

```{r model}
treat_model <- MASS::polr(COVIDseverity~has_cf + NC + HFNC,
                          data = treatments_df, Hess = TRUE)
summary(treat_model)
```

P-values are not shown in the summary table. Therefore, will create a a table to 
highlight p values and coefficients together 

```{r grouped coefficients}
ctable1 <- coef(summary(treat_model))

## calculate and store p values
p <- round(pnorm(abs(ctable1[, "t value"]), lower.tail = FALSE) * 2, 3)

## combined table
(ctable1 <- cbind(ctable1, "p value" = p))
```


