---
title: "Child Patients with Covid-19 and Cystic Fibrosis"
author: "Alfonso Vieyra, Brandon Huett, Allison Cher, Brittnie Villasenor, Ernesto Collazo"
date: '2022-06-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(scales)

# Need to always load the package though
library(epitools)
library(rmeta)
library(pROC)
library(nnet)
```

## Ordinal multinomial logistic Model with non grouped moorbidities 

Will first create a sub data frame consisting only of patients who reported 
their first visit and columns consisting of the morbidities, has_cf and 
COVIDseverity.

```{r DataFrame}
unique_patient_df <- 
  d[order(d$servicedate),] %>%
  distinct(personid, .keep_all = TRUE)

morb_df <- unique_patient_df %>%
  select(has_cf, COVIDseverity, starts_with('comorb'))
colnames(morb_df) <- c('has_cf', 'COVIDseverity', 'bronchiectasis',
                        'pulmonary_disease', 'asthma', 'nasal_polyps',
                        'hemoptysis', 'pneumothorax', 'resp_failure', 
                        'malnutrition', 'other_nutri_deficiency', 
                        'type1_diabetes', 'type2_diabetes', 'obesity', 
                        'liver_disease', 'other_GI_notLiver', 
                        'essential_hypertension', 'hypertensive_heart_disease',
                        'chronic_kidney_disease', 'nicotine_dependence', 
                        'heart_failure', 'ischemic_heart_disease', 
                        'lung_transplant')
```

Implementation of the model here:

```{r full model}
full_model <- MASS::polr(COVIDseverity~., data = morb_df, Hess = TRUE)
full_sum <- summary(full_model)
```

P-values are not shown in the summary table. Therefore, will create a a table to 
highlight p values and coefficients together 

```{r non group coefficients}
ctable1 <- coef(summary(full_model))

## calculate and store p values
p <- round(pnorm(abs(ctable1[, "t value"]), lower.tail = FALSE) * 2, 3)

## combined table
(ctable1 <- cbind(ctable1, "p value" = p))
```
We can see that the following predictors have no significance in predicting lvls
of severity: has_cf, pulmonary disease, type 2 diabetes, hypertensive heart 
disease and ishemic heart disease. 

The fact that has_cf has a small p value, tells us that there is no 
discernible way to determine the odds of a patient having levels 0, 1 or 2 covid
severity between non-cf and cf patients. 

## Ordinal multinomial logistic Model with grouped moorbidities 

using the morb_df object and instantiating a new object with comorbidities 
grouped together.

```{r grouping morbidities}
morb_grouped <- morb_df %>% 
  mutate(
    respiratory_disease = case_when(
      bronchiectasis == 1|
        pulmonary_disease == 1|
        asthma == 1|
        nasal_polyps == 1|
        hemoptysis == 1|
        pneumothorax == 1|
        resp_failure == 1|
        lung_transplant == 1~1,
      bronchiectasis == 0&
        pulmonary_disease == 0&
        asthma == 0&
        nasal_polyps == 0&
        hemoptysis == 0&
        pneumothorax == 0&
        resp_failure == 0&
        lung_transplant == 0~0),
    
    nutri_cond = case_when(
      malnutrition == 1|
        other_nutri_deficiency == 1|
        obesity == 1~1,
      malnutrition == 0&
        other_nutri_deficiency == 0&
        obesity == 0~0),
    
    diabetes = case_when(
      type1_diabetes == 1|
        type2_diabetes == 1~1,
      type1_diabetes == 0&
        type2_diabetes == 0~0),
    
    cardiac_dis = case_when(
      heart_failure == 1|
        hypertensive_heart_disease == 1|
        ischemic_heart_disease == 1|
        essential_hypertension == 1~1,
      heart_failure == 0&
        hypertensive_heart_disease == 0&
        ischemic_heart_disease == 0&
        essential_hypertension == 0~0),
      
      abdominal_cond = case_when(
        liver_disease == 1|
          other_GI_notLiver == 1|
          chronic_kidney_disease == 1~1,
        liver_disease == 0&
          other_GI_notLiver == 0&
          chronic_kidney_disease == 0~0),
    
    nicotine_dep = case_when(
      nicotine_dependence == 1~1,
      nicotine_dependence == 0~0)
    )

morb_grouped <- morb_grouped %>%
  select(1:2, 24:29)
```

Employing model here:

```{r grouped model}
group_model <- MASS::polr(COVIDseverity~., data = morb_grouped, Hess = TRUE)
summary(group_model)
```

```{r grouped coefficients}
ctable2 <- coef(summary(group_model))

## calculate and store p values
p <- round(pnorm(abs(ctable2[, "t value"]), lower.tail = FALSE) * 2, 3)

## combined table
(ctable2 <- cbind(ctable2, "p value" = p))
```


We can see that testing for groups we come up with p-values smaller across the 
board with the exception of having cf or not. 

## Modeling for the number of Comorbidities Per Patient 

Creating the data frame that consists of the number of morbidities per patient
(row) with has_cf and COVIDseverity.

```{r morb_perP df}
morbs <- morb_df %>%
  select(3:23)

morbs[] <- lapply(morbs, function(x) as.numeric(as.character(x)))

morbs <- morbs %>%
     mutate(total_morbs = rowSums(.),
            has_cf = morb_df$has_cf,
  COVIDseverity = morb_df$COVIDseverity) %>%
  select(has_cf, COVIDseverity, total_morbs)
```

Implementation of the model is here: 

COVIDseverity = B0 + B1(Has_cf) + B2(total_comorbidities)

```{r model}
perP_model <- MASS::polr(COVIDseverity~has_cf + total_morbs, 
                         data = morbs, Hess = TRUE)
summary(perP_model)
```

Creating table to calculate p-values and show all estimates.

```{r coefficients3}
ctable3 <- coef(summary(perP_model))

## calculate and store p values
p <- round(pnorm(abs(ctable3[, "t value"]), lower.tail = FALSE) * 2, 3)

## combined table
(ctable3 <- cbind(ctable3, "p value" = p))
```

We can see that the p value for has_cf is high. This suggests further that 
having an illness like cystic fibrosis has no increased likelihood or lack
thereof of ones covid severity symptoms. However, interestingly enough, we can
see that there is significance in the number of morbidities per patient and 
there increased likelihood at contracting severe symptoms. This is odd since 
patients who have cystic fibrosis have a greater proportion of its sample with 
pre existing conditions. 

I did conduct a model including an interaction term between having cystic 
fibrosis and the number of comorbidities present in a patient. However, both
have_cf and the interaction term yielded large p-values and again restating its 
lack of significance in the presence of prexisting conditions. 

## Conclusion

We can conclude that having cystic fibrosis under the presence of comorbidities 
as additional predictors alone, does not present any increased risk at 
contracting severe symptoms versus patients who do not have cystic fibrosis. In
other words, the odds of contracting severe symptoms is independent of having
cystic fibrosis. May look at possible interactions to determine any difference. 