## Part 3

```{r}
library(tidyverse)
library(scales)
library(lubridate)
library(ggplot2)
library(aod)
library(mlbench)
```

```{r}
filter(d, has_cf == 1)
sub_wcf <- filter(d, has_cf == 1)
sub_wocf <- filter(d, has_cf == 0)
```

```{r}
morb_df <- d %>%
  select(has_cf, COVIDseverity, starts_with('comorb'))

colnames(morb_df) <- c('has_cf', 'COVIDseverity', 'bronchiectasis',
                        'pulmonary_disease', 'asthma', 'nasal_polyps',
                        'hemoptysis', 'pneumothorax', 'resp_failure', 
                        'malnutrition', 'other_nutri_deficiency', 
                        'type1_diabetes', 'type2_diabetes', 'obesity', 
                        'liver_disease', 'other_GI_notLiver', 
                        'essential_hypertension', 'hypertensive_heart_disease',
                        'chronic_kidney_disease', 'nicotine_dependence', 
                        'heart_failure', 'ischemic_heart_disease', 
                        'lung_transplant')
#table of non cystic fibrosis patients
morb_noCF <- morb_df  %>%
  filter(has_cf == 0)
#table of cystic fibrosis patients
morb_CF <- morb_df %>%
  filter(has_cf == 1)
```

```{r}
#created altered data frame of morb_df to have severity to two levels
morb_df_01 <- morb_df
morb_df_01['COVIDseverity'][morb_df_01['COVIDseverity'] == 2] <- 1

morb_df_01$COVIDseverity <- factor(morb_df_01$COVIDseverity, c(0,1))
```

```{r group morbids}
morb_grouped <- morb_df_01 %>% 
  mutate(
    respiratory_disease = case_when(
      bronchiectasis == 1|
        pulmonary_disease == 1|
        asthma == 1|
        nasal_polyps == 1|
        hemoptysis == 1|
        pneumothorax == 1|
        resp_failure == 1|
        lung_transplant == 1~1,
      bronchiectasis == 0&
        pulmonary_disease == 0&
        asthma == 0&
        nasal_polyps == 0&
        hemoptysis == 0&
        pneumothorax == 0&
        resp_failure == 0&
        lung_transplant == 0~0),
    
    nutri_cond = case_when(
      malnutrition == 1|
        other_nutri_deficiency == 1|
        obesity == 1~1,
      malnutrition == 0&
        other_nutri_deficiency == 0&
        obesity == 0~0),
    
    diabetes = case_when(
      type1_diabetes == 1|
        type2_diabetes == 1~1,
      type1_diabetes == 0&
        type2_diabetes == 0~0),
    
    cardiac_dis = case_when(
      heart_failure == 1|
        hypertensive_heart_disease == 1|
        ischemic_heart_disease == 1|
        essential_hypertension == 1~1,
      heart_failure == 0&
        hypertensive_heart_disease == 0&
        ischemic_heart_disease == 0&
        essential_hypertension == 0~0),
      
      abdominal_cond = case_when(
        liver_disease == 1|
          other_GI_notLiver == 1|
          chronic_kidney_disease == 1~1,
        liver_disease == 0&
          other_GI_notLiver == 0&
          chronic_kidney_disease == 0~0),
    
    nicotine_dep = case_when(
      nicotine_dependence == 1~1,
      nicotine_dependence == 0~0)
    )

morb_grouped <- morb_grouped %>%
  select(1:2, 24:29)
```

```{r glm for each morb}
set.seed(1234)
summary(glm(COVIDseverity~respiratory_disease+has_cf, family = binomial(link='logit'), 
            data = morb_grouped))

summary(glm(COVIDseverity~nutri_cond+has_cf, family = binomial(link='logit'), 
            data = morb_grouped))

summary(glm(COVIDseverity~diabetes+has_cf, family = binomial(link='logit'), 
            data = morb_grouped))

summary(glm(COVIDseverity~cardiac_dis+has_cf, family = binomial(link='logit'), 
            data = morb_grouped))

summary(glm(COVIDseverity~abdominal_cond+has_cf, family = binomial(link='logit'), 
            data = morb_grouped))

summary(glm(COVIDseverity~nicotine_dep+has_cf, family = binomial(link='logit'), 
            data = morb_grouped))

```

```{r glm for all morb}
set.seed(1234)
summary(glm(COVIDseverity~., family = binomial(link='logit'), 
            data = morb_grouped))
```

```{r sub_df of morb_grouops without diabetes}
no_diab_morb <- morb_grouped %>%
  select(-diabetes)
```

```{r logit model sub_df}
model_morb <-
glm(COVIDseverity~., family = binomial(link='logit'), 
            data = no_diab_morb)
summary(model_morb)
```

```{r}
a=seq(1,925952,1)
b=sample(a,740762,replace = F)
no_diab_morb_train=no_diab_morb[b,]
no_diab_morb_test=no_diab_morb[-b,]
no_diab_morb_train %>% filter(has_cf == 1) %>% count()
```

```{r}
no_diab_fitted <- glm(COVIDseverity ~ ., family = binomial(link='logit'), data = no_diab_morb_train)
```

```{r}
no_diab_predict_fit = no_diab_fitted %>% predict(no_diab_morb_test, type="response")
```

```{r}
predicted_severity2 <- ifelse(no_diab_predict_fit > 0.5, 1,0)
```

```{r}
counter=0
for(i in 1:185190)
{
  if(predicted_severity2[i] != no_diab_morb_test$COVIDseverity[i])
  {counter=counter+1}
}
counter
misclassification.rate=counter/185190
misclassification.rate
```
```

```{r prop comp}
# Need to always load the package though
library(epitools)
library(rmeta)
library(pROC)
library(nnet)

#####################################################
#####################################################
# This is the prop.comp() function we will use
#######
prop.comp <- function( x, estimate="all", conf.level=.95, transpose=FALSE ){
	if( transpose ) x <- t(x)
	rslt <- vector( "list", length=3 )
	names( rslt ) <- c( "riskdiff", "riskratio", "oddsratio" )
	diff.rslt <- suppressWarnings(prop.test( x, conf.level=conf.level ))
	rslt[[1]] <- rslt[[2]] <- rslt[[3]] <- epitab( x, method="riskratio", pvalue="chi2", conf.level=conf.level )$tab
	colnames( rslt[[1]] )[5] <- "riskdiff"
	rslt[[1]][,5] <- c(0,diff(rev(diff.rslt$estimate)))
	rslt[[1]][2,6:7] <- diff.rslt$conf.int
	colnames( rslt[[3]] )[5] <- "oddsratio"
	rslt[[3]][,5:8] <- suppressWarnings(epitab( x, method="oddsratio", pvalue="chi2", conf.level=conf.level )$tab[,5:8])
	if(is.null(names(dimnames(x)))){
		for(i in 1:3){
			colnames(rslt[[i]])[c(1,3)] <- c("Outcome=0", "Outcome=1")
			rownames(rslt[[i]]) <- c("Group=1", "Group=2")
			}
	}
	if( is.element( estimate, c("all", "oddsratio") ) ){ 
		if(is.null(names(dimnames(x)))){
			warning( "Estimated probabilities represent Pr[ Outcome | Group ]. For estimates of 
			Pr[ Group | Outcome ], change the value of 'transpose'.")
		}
		else
			warning( paste("Estimated probabilities represent Pr[", names(dimnames(x))[2], 
			"|",names(dimnames(x))[1], "]. For estimates of 
			Pr[", names(dimnames(x))[1], "|",names(dimnames(x))[2], "], change the value of 'transpose'.") )
		}
	if( estimate == "riskdiff" ) return(rslt[[1]])
	else if( estimate == "riskratio" ) return(rslt[[2]])
	else if( estimate == "oddsratio" ) return(rslt[[3]])
	else return(rslt)
}
```

```{r}
#cont_bronc is the data frame which features severity lvls and comorb 
#bronchietysis. Morb_df is the data frame featuring has_cf, severity and all 
#morbidities

cont_bronc <- morb_df_01 %>%
  select(COVIDseverity, bronchiectasis)

# changed all patients whose severity was 2 to 1. 
cont_bronc['COVIDseverity'][cont_bronc['COVIDseverity'] == 2] <- 1

#refactored into 2 lvls: 0 and 1 as opposed to three lvls so that prop.comp
#can be used
cont_bronc$COVIDseverity <- factor(cont_bronc$COVIDseverity, c(0,1))

#converted to contingency table
cont_bronc <-cont_bronc %>%
  table()

#use prop comp to come up with risk diff, risk ratio and odds ratio
prop.comp(cont_bronc)
# this will present the risk ratio, difference and etc
```

```{r}
sub_wcf %>%
  select(COVIDseverity, comorb_bronchiectasis_J47) %>%
  table()

```

```{r}
sub_wocf %>%
  select(COVIDseverity, comorb_bronchiectasis_J47) %>%
  table()
  
```

```{r}
hr <- filter(d, heartrate <= 220)
ppp <- ggplot(hr, aes(x = COVIDseverity, y = heartrate, fill = has_cf)) +
  geom_boxplot() + 
  ggtitle("Covid 19 Severity Based on Patients Heartrate")
  scale_fill_manual(labels = c("Covid 19 Severity", "Heartrate of the Patient"))
ppp <- ppp + guides(fill=guide_legend(title="CF Indicator"))
ppp
```

```{r}
p <- ggplot(sub_wcf, aes(x = COVIDseverity, fill = race)) +
  geom_bar() + 
  labs(x = "Covid 19 Severity",
       y = "Number of Patients",
       title = "Covid 19 Severity Within each Race of Patients with CF")
p <- p + guides(fill=guide_legend(title="Race"))
p
```

```{r}
pp <- ggplot(sub_wocf, aes(x = COVIDseverity, fill = race)) +
  geom_bar() + 
  labs(x = "Covid 19 Severity",
       y = "Number of Patients",
       title = "Covid 19 Severity Within each Race of Patients without CF")
pp <- pp + guides(fill=guide_legend(title="Race"))
pp
```

```{r}
# example of contigency table
table(d$COVIDseverity, d$INOTROPES)
```

```{r}
# patients with cf
sub_df %>%
  group_by(COVIDseverity, comorb_bronchiectasis_J47) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_bronchiectasis_J47) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
# patients with cf
sub_df %>%
  group_by(COVIDseverity, comorb_copd_J40_J44) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_copd_J40_J44) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
# should look at asthma
# patients with cf
sub_df %>%
  group_by(COVIDseverity, comorb_asthma_J45) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_asthma_J45) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
# patients with cf
sub_df %>%
  group_by(COVIDseverity, comorb_nasal_polyps_J33) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_nasal_polyps_J33) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_hemoptysisR042) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_hemoptysisR042) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_pneumothorax_J93) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_pneumothorax_J93) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
# should look at resp failure
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_resp_failure_J96) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_resp_failure_J96) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_pneumothorax_J93) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_pneumothorax_J93) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_malnutrition) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_malnutrition) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_other_nutritional_deficiencies_E50_E64) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_other_nutritional_deficiencies_E50_E64) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_typeI_diabetes_E10) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_typeI_diabetes_E10) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_typeII_diabetes_E11) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_typeII_diabetes_E11) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_obesity_overweight_E66) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_obesity_overweight_E66) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_liver_disease_K70_K77) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_liver_disease_K70_K77) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
# this crazy to look at
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_other_GI_notLiver_K_excludesK70K77) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_other_GI_notLiver_K_excludesK70K77) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_essential_hypertension_I10) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_essential_hypertension_I10) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_hypertensive_heart_disease_I11) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_hypertensive_heart_disease_I11) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_chronic_kidney_disease_N18) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_chronic_kidney_disease_N18) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_nicotine_dependence_F17) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_nicotine_dependence_F17) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_heart_failure_I50) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_heart_failure_I50) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_ischemic_heart_disease_I20_I25) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_ischemic_heart_disease_I20_I25) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_lung_transplant_Z942) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_lung_transplant_Z942) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}

```
## Part 4

```{r}
con_tab <- d %>% 
  select(COVIDseverity, has_cf, age_at_encounter)
```


```{r}
#contingency table, compares non-cf covid severity with race
non_cf %>% 
  group_by(COVIDseverity, encountertype) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(N / sum(N), 3)) %>% 
  knitr::kable()
```

```{r}
#contingency table, compares cf covid severity with encountertype
cf %>% 
  group_by(COVIDseverity, encountertype) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(N / sum(N), 3)) %>% 
  knitr::kable()
```

```{r}
#contingency table, compares cf covid severity with encountertype
non_cf %>% 
  group_by(COVIDseverity, encountertype) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(N / sum(N), 3)) %>% 
  knitr::kable()
```

```{r}
#contingency table, compares covid severity between cf and non-cf
con_tab %>% 
  group_by(has_cf, COVIDseverity) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round((N / sum(N)), 3)) %>% 
  knitr::kable()
```

```{r}
#contingency table, compares covid severity between cf and non-cf (shows percentage)
con_tab %>% 
#contingency table, compares covid severity between cf and non-cf
ratetable <- con_tab %>% 
  group_by(has_cf, COVIDseverity) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
ratetable <- con_tab %>% 
  group_by(has_cf, COVIDseverity) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round((N / sum(N)), 3))
```


```{r}
#bar plot, covid severity for cf patients
ggplot(ratetable, aes(x = has_cf,
                      y = Ratio,
                      fill = COVIDseverity), position = "dodge") +
    labs(x = "Patients Without CF", 
       y = "Rate", 
       title = "COVID Severity Among Children With/Without Cystic Fibrosis") +
  geom_bar(stat = "identity")
```

```{r}
library(ggplot2)
theme_set(theme_classic())

ratetable <-
```


```{r}
#bar plot, covid severity for non-cf patients
ggplot(non_cf) +
  geom_bar(mapping = aes(x=has_cf, fill = COVIDseverity), 
           position = 'dodge') +
  scale_y_continuous(labels = comma) +
  labs(x = "Patients Without CF", 
       y = "Count", 
       title = "COVID Severity Among Children Without Cystic Fibrosis")
```

```{r}
ggplot(con_tab) +
  geom_bar(mapping = aes(x=has_cf, fill = COVIDseverity), 
           position = 'dodge') +
  coord_cartesian(ylim = c(0, 60000)) +
  scale_y_continuous(labels = comma)
```

Logistic Regression
---------------------
```{r}
library(aod)
library(ggplot2)
```

```{r}
covsev01 <- d
```

```{r}
covsev01 <-
  d["COVIDseverity"][d["COVIDseverity"] == 2] <- 1
```

```{r}
covid_data2 <- d %>% 
  select(COVIDseverity, has_cf, age_at_encounter)
```

```{r}
covid_data2["COVIDseverity"][covid_data2["COVIDseverity"] == 2] <- 1
```

```{r}
model <- glm(COVIDseverity ~., family = binomial(link='logit'),
             data = covid_data2)
summary(model)
```

```{r}
confint(model)
```

```{r}
confint.default(model)
```

```{r}
exp(coef(model))
```
```{r}
exp(cbind(OR = coef(model), confint(model)))
```

```{r}
wald.test(b = coef(model), Sigma = vcov(model), Terms = 1:2)
```

```{r}
ggplot(covid_data2, aes(x = has_cf, y = COVIDseverity)) + geom_ribbon(aes(ymin = LL,
    ymax = UL, fill = rank), alpha = 0.2) + geom_line(aes(colour = rank),
    size = 1)
```
