---
title: "Child Patients with Covid-19 and Cystic Fibrosis"
author: "Alfonso Vieyra"
date: '2022-06-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(scales)
# Need to always load the package though
library(epitools)
library(rmeta)
library(pROC)
library(nnet)
```

```{r prop comp}
# This is the prop.comp() function we will use
#######
prop.comp <- function( x, estimate="all", conf.level=.95, transpose=FALSE ){
	if( transpose ) x <- t(x)
	rslt <- vector( "list", length=3 )
	names( rslt ) <- c( "riskdiff", "riskratio", "oddsratio" )
	diff.rslt <- suppressWarnings(prop.test( x, conf.level=conf.level ))
	rslt[[1]] <- rslt[[2]] <- rslt[[3]] <- epitab( x, method="riskratio", pvalue="chi2", conf.level=conf.level )$tab
	colnames( rslt[[1]] )[5] <- "riskdiff"
	rslt[[1]][,5] <- c(0,diff(rev(diff.rslt$estimate)))
	rslt[[1]][2,6:7] <- diff.rslt$conf.int
	colnames( rslt[[3]] )[5] <- "oddsratio"
	rslt[[3]][,5:8] <- suppressWarnings(epitab( x, method="oddsratio", pvalue="chi2", conf.level=conf.level )$tab[,5:8])
	if(is.null(names(dimnames(x)))){
		for(i in 1:3){
			colnames(rslt[[i]])[c(1,3)] <- c("Outcome=0", "Outcome=1")
			rownames(rslt[[i]]) <- c("Group=1", "Group=2")
			}
	}
	if( is.element( estimate, c("all", "oddsratio") ) ){ 
		if(is.null(names(dimnames(x)))){
			warning( "Estimated probabilities represent Pr[ Outcome | Group ]. For estimates of 
			Pr[ Group | Outcome ], change the value of 'transpose'.")
		}
		else
			warning( paste("Estimated probabilities represent Pr[", names(dimnames(x))[2], 
			"|",names(dimnames(x))[1], "]. For estimates of 
			Pr[", names(dimnames(x))[1], "|",names(dimnames(x))[2], "], change the value of 'transpose'.") )
		}
	if( estimate == "riskdiff" ) return(rslt[[1]])
	else if( estimate == "riskratio" ) return(rslt[[2]])
	else if( estimate == "oddsratio" ) return(rslt[[3]])
	else return(rslt)
}
```

## Part 1

```{r morbidity Tables}
#table with severity, cf and comorbidities NOTE: Changed names for readability
morb_df <- d %>%
  select(has_cf, COVIDseverity, starts_with('comorb'))


colnames(morb_df) <- c('has_cf', 'COVIDseverity', 'bronchiectasis',
                        'pulmonary_disease', 'asthma', 'nasal_polyps',
                        'hemoptysis', 'pneumothorax', 'resp_failure', 
                        'malnutrition', 'other_nutri_deficiency', 
                        'type1_diabetes', 'type2_diabetes', 'obesity', 
                        'liver_disease', 'other_GI_notLiver', 
                        'essential_hypertension', 'hypertensive_heart_disease',
                        'chronic_kidney_disease', 'nicotine_dependence', 
                        'heart_failure', 'ischemic_heart_disease', 
                        'lung_transplant')

#table of non cystic fibrosis patients
morb_noCF <- morb_df  %>%
  filter(has_cf == 0)

#table of cystic fibrosis patients
morb_CF <- morb_df %>%
  filter(has_cf == 1)
```

```{r funcy}
#created a function to calculate the proportions in a 1D vector
prop1 <- function(arr, denom) {
  count = 1
  for (x in arr)
  {
    arr[count] = round(x / denom, 2)
    count = count + 1
  }
  return(arr)
}
```

```{r Severity Levels noCF}

#Tables of severity symptoms level 0
severity0_noCF <- morb_noCF %>%
  filter(COVIDseverity == 0) %>%
  select(-has_cf, -COVIDseverity)

indx0 <- sapply(severity0_noCF, is.factor)
severity0_noCF[indx0] <- lapply(severity0_noCF[indx0], 
                           function(x) as.numeric(as.character(x)))

#Tables of severity symptoms level 1
severity1_noCF <- morb_noCF %>%
  filter(COVIDseverity == 1) %>%
  select(-has_cf, -COVIDseverity)

indx1 <- sapply(severity1_noCF, is.factor)
severity1_noCF[indx1] <- lapply(severity1_noCF[indx1], 
                           function(x) as.numeric(as.character(x)))

#Tables of severity symptoms level 2 
severity2_noCF <- morb_noCF %>%
  filter(COVIDseverity == 2) %>%
  select(-has_cf, -COVIDseverity)

indx2 <- sapply(severity2_noCF, is.factor)
severity2_noCF[indx2] <- lapply(severity2_noCF[indx2], 
                           function(x) as.numeric(as.character(x)))

#Vectorized severities of counts
asymptomatic <- colSums(severity0_noCF)
mild_symptoms <- colSums(severity1_noCF)
severe_symptoms <- colSums(severity2_noCF)

#created new df based on counts of morbidities across severity lvls 
library(reshape2)

sev_lvls_noCF <- rbind(asymptomatic, mild_symptoms, severe_symptoms)
sev_lvls_noCF[1, ] <- prop1(sev_lvls_noCF[1,], 296236)
sev_lvls_noCF[2, ] <- prop1(sev_lvls_noCF[2,], 34015)
sev_lvls_noCF[3, ] <- prop1(sev_lvls_noCF[3,], 20743)

melted_sev_lvls_noCF <- melt(sev_lvls_noCF)
colnames(melted_sev_lvls_noCF) <- c('severity_lvl', 'morbidity','proportion')
```


```{r visual non-CF}
options(scipen = 999)
ggplot(melted_sev_lvls_noCF, aes(x=morbidity, y = proportion, fill = proportion)) +
  geom_bar(stat = 'identity') +
  coord_flip() +
  labs( x = 'Morbidities', y = 'Total Reported Conditions') +
  ggtitle('Dominant Conditions in Non-CF patients with Covid 19') +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~severity_lvl) + 
  theme(axis.text.x = element_text(angle=45))
```

```{r Severity Levels CF}
#Tables of severity symptoms level 0
severity0_CF <- morb_CF %>%
  filter(COVIDseverity == 0) %>%
  select(-has_cf, -COVIDseverity)

indx0_cf <- sapply(severity0_CF, is.factor)
severity0_CF[indx0_cf] <- lapply(severity0_CF[indx0_cf], 
                           function(x) as.numeric(as.character(x)))

#Tables of severity symptoms level 1
severity1_CF <- morb_CF %>%
  filter(COVIDseverity == 1) %>%
  select(-has_cf, -COVIDseverity)

indx1_cf <- sapply(severity1_CF, is.factor)
severity1_CF[indx1_cf] <- lapply(severity1_CF[indx1_cf], 
                           function(x) as.numeric(as.character(x)))

#Tables of severity symptoms level 2 
severity2_CF <- morb_CF %>%
  filter(COVIDseverity == 2) %>%
  select(-has_cf, -COVIDseverity)

indx2_cf <- sapply(severity2_CF, is.factor)
severity2_CF[indx2_cf] <- lapply(severity2_CF[indx2_cf], 
                           function(x) as.numeric(as.character(x)))

#Vectorized severities of counts
asympt_mild <- colSums(severity0_CF)
moderate_symptoms <- colSums(severity1_CF)
severe_symptoms <- colSums(severity2_CF)

#created new df based on counts of morbidities across severity lvls 
sev_lvls_CF <- rbind(asympt_mild, moderate_symptoms, severe_symptoms)
sev_lvls_CF[1,] <- prop1(sev_lvls_CF[1,], 8454)
sev_lvls_CF[2,] <- prop1(sev_lvls_CF[2,], 1216)
sev_lvls_CF[3,] <- prop1(sev_lvls_CF[3,], 646)

melted_sev_lvls_CF <- melt(sev_lvls_CF)
colnames(melted_sev_lvls_CF) <- c('severity_lvl', 'morbidity','proportion')
```

```{r visual CF}
ggplot(melted_sev_lvls_CF, aes(x=morbidity, y = proportion, fill = proportion)) +
  geom_bar(stat = 'identity') +
  coord_flip()+
  labs( x = 'Morbidities', y = 'Proportion of Conditions') +
  ggtitle('Dominant Conditions in CF patients with Covid 19') +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~severity_lvl) + 
  theme(axis.text.x = element_text(angle=45))
```

```{r cf non cf}
has_CF <- morb_CF %>%
  select(-has_cf, -COVIDseverity)

indx_hasCF <- sapply(has_CF, is.factor)
has_CF[indx_hasCF] <- lapply(has_CF[indx_hasCF], 
                           function(x) as.numeric(as.character(x)))

has_noCF <- morb_noCF %>%
  select(-has_cf, -COVIDseverity)

indx_hasNoCF <- sapply(has_noCF, is.factor)
has_noCF[indx_hasNoCF] <- lapply(has_noCF[indx_hasNoCF], 
                           function(x) as.numeric(as.character(x)))

cyst_fibrosis_patients <- colSums(has_CF)
noncyst_fibrosis_patients <- colSums(has_noCF)

cf_lvls <- rbind(cyst_fibrosis_patients, noncyst_fibrosis_patients)
cf_lvls[1,] <- prop1(cf_lvls[1,], 10316)
cf_lvls[2,] <- prop1(cf_lvls[2,], 350994)

melted_cf_lvls <- melt(cf_lvls)
colnames(melted_cf_lvls) <- c('has_CF', 'morbidity','proportion')
```

```{r visuals cf lvls}
ggplot(melted_cf_lvls, aes(x=morbidity, y = proportion, fill = proportion)) +
  geom_bar(stat = 'identity') +
  coord_flip()+
  labs( x = 'Morbidities', y = 'Proportion of Conditions') +
  ggtitle('Dominant Conditions in CF and non-CF patients with Covid 19') +
  theme(plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~has_CF) + 
  theme(axis.text.x = element_text(angle=45))
```

```{r # of morbidities in patients}
#number of mordidities in each patient
num_morb_perCF <- table(rowSums(has_CF))
num_morb_perNonCF <- table(rowSums(has_noCF))

#Assign proportions for the number of comorbidities per patient
num_morb_perCF <- prop1(num_morb_perCF, 9337)
num_morb_perNonCF <- prop1(num_morb_perNonCF, 916615)
```

```{r logit}
summary(glm(has_cf~comorb_bronchiectasis_J47, family = binomial(link='identity'), data = d))
```


## Part 2
```{r}
library(lubridate)
```

##create subset of just non cf patients
```{r}
non_cf <- d %>% filter(has_cf == 0)
View(non_cf)
```

##type of treatment determines covidseverity (0,1,2)
```{r}
severity <- d %>% select(personid, COVIDseverity, ECMO, HFNC, NC, bipapCpapNonInvasive, invasiveMechVentilation)
view(severity)
```

##counts of severity level based on cf and non cf patients
```{r}
ggplot(d, aes(x=COVIDseverity, fill = has_cf)) + geom_bar(position = "dodge")
```

##create subset that includes admission and discharge date/time for all patients
```{r}
duration <- d %>% select(servicedate, dischargedate)
```

##get rid of the letter "T" and "Z" in the service date column.
```{r}
duration$servicedate <- gsub("T"," ", duration$servicedate)
duration$servicedate <- gsub("Z"," ", duration$servicedate)
```

##Get rid of the letter "T" and "Z" in the discharge date column.
```{r}
duration$dischargedate <- gsub("T", " ", duration$dischargedate)
duration$dischargedate <- gsub("Z", " ", duration$dischargedate)
```

##Get rid of ".000" in both columns.
```{r}
duration$servicedate <- gsub(".000"," ", duration$servicedate)
duration$dischargedate <- gsub(".000"," ", duration$dischargedate)
```


##Transform dates stored as character or numeric vectors to POSIXct objects. 
```{r}
duration$servicedate <- ymd_hms(duration$servicedate)
duration$dischargedate <- ymd_hms(duration$dischargedate)
```

##Subtract discharge date by service date to find length of hospital stay in seconds.
```{r}
duration$diffsecs <- duration$dischargedate-duration$servicedate
```

##Calculate the length of hospital stay in minutes. Create column for minutes.
```{r}
duration$diffmins=(duration$dischargedate-duration$servicedate)/60
duration$diffmins <- gsub("secs","mins", duration$diffmins)
```

##Calulate the length of hospital stay in hours
```{r}
duration$diffhrs=(duration$dischargedate-duration$servicedate)/3600
```

##Attach "has_cf" column from d dataset to duration subset
```{r}
duration$has_cf = d$has_cf
```

##Filter the length of time for cf and non cf patients.
```{r}
time_non_cf <- duration %>% filter(has_cf == 0)
time_has_cf <- duration %>% filter(has_cf == 1)
```

##non_cf: convert diffmins from character to numeric variable. convert diffhours to numeric.
```{r}
time_non_cf <- time_non_cf %>% mutate(diffmins=as.numeric(diffmins))
time_non_cf <- time_non_cf %>% mutate(diffhrs=as.numeric(diffhrs))
```

##has_cf: convert diffmins from character to numeric variable. convert diffhours to numeric.
```{r}
time_has_cf <- time_has_cf %>% mutate(diffmins = as.numeric(diffmins))
time_has_cf <- time_has_cf %>% mutate(diffhrs = as.numeric(diffhrs))
```

##get rid of NA values in time_has_cf subset.
```{r}
time_has_cf <- na.omit(time_has_cf)
```

##make any negative values in time_non_cf subset to NA. Get rid of NA values.
```{r}
time_non_cf[time_non_cf < 0] <- NA
time_non_cf <- na.omit(time_non_cf)
```

##histogram of hours cf children spent at the hospital. does not include outliers. hours limited to 700 for visibility.
```{r}
plot_has_cf_lim <- ggplot(time_has_cf, aes(x=diffhrs)) + geom_histogram(bins=200) + coord_cartesian(xlim = c(0,700)) + ggtitle("Hours CF Children Spent in the Hospital") +theme(plot.title = element_text(hjust = 0.5))

```


##histogram of hours cf children spent at the hospital. includes outliers 
```{r}
plot_has_cf_total <- ggplot(time_has_cf, aes(x=diffhrs)) + geom_histogram() + ggtitle("Hours CF Children Spent in the Hospital") +theme(plot.title = element_text(hjust = 0.5))
```

##histogram of hours non cf children spent at the hospital. includes outliers.
```{r}
plot_non_cf_total <- ggplot(time_non_cf, aes(x=diffhrs)) + geom_histogram() + ggtitle("Hours Non-CF Children Spent in the Hospital") +theme(plot.title = element_text(hjust = 0.5))
```

##histogram of hours non cf children spent at the hospital. does not include outliers. hours limited to 2500 for visibility.
```{r}
plot_non_cf_lim <-ggplot(time_non_cf, aes(x=diffhrs)) + geom_histogram() + coord_cartesian(xlim = c(0,2500)) + ggtitle("Hours Non-CF Children Spent in the Hospital") +theme(plot.title = element_text(hjust = 0.5))
```


## Part 3
```{r}
library(tidyverse)
filter(d, has_cf == 1)
sub_df <- filter(d, has_cf == 1)
nonsub_df <- filter(d, has_cf == 0)
```

```{r}
d %>%
  select(COVIDseverity, comorb_lung_transplant_Z942) %>%
  table()

```

```{r}
sub_df %>%
  select(COVIDseverity, comorb_lung_transplant_Z942) %>%
  table()
  
```

```{r}
hr <- filter(d, heartrate <= 220)
ppp <- ggplot(hr, aes(x = COVIDseverity, y = heartrate, fill = has_cf)) +
  geom_boxplot() + 
  ggtitle("Covid 19 Severity Based on Patients Heartrate")
  scale_fill_manual(labels = c("Covid 19 Severity", "Heartrate of the Patient"))
ppp <- ppp + guides(fill=guide_legend(title="CF Indicator"))
ppp
```

```{r}
p <- ggplot(sub_df, aes(x = COVIDseverity, fill = race)) +
  geom_bar() + 
  labs(x = "Covid 19 Severity",
       y = "Number of Patients",
       title = "Covid 19 Severity Within each Race of Patients with CF")
p <- p + guides(fill=guide_legend(title="Race"))
p
```

```{r}
pp <- ggplot(nonsub_df, aes(x = COVIDseverity, fill = race)) +
  geom_bar() + 
  labs(x = "Covid 19 Severity",
       y = "Number of Patients",
       title = "Covid 19 Severity Within each Race of Patients without CF")
pp <- pp + guides(fill=guide_legend(title="Race"))
pp
```

```{r}
# example of contigency table
table(d$COVIDseverity, d$INOTROPES)
```

```{r}
# patients with cf
sub_df %>%
  group_by(COVIDseverity, comorb_bronchiectasis_J47) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_bronchiectasis_J47) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
# patients with cf
sub_df %>%
  group_by(COVIDseverity, comorb_copd_J40_J44) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_copd_J40_J44) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
# should look at asthma
# patients with cf
sub_df %>%
  group_by(COVIDseverity, comorb_asthma_J45) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_asthma_J45) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
# patients with cf
sub_df %>%
  group_by(COVIDseverity, comorb_nasal_polyps_J33) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_nasal_polyps_J33) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_hemoptysisR042) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_hemoptysisR042) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_pneumothorax_J93) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_pneumothorax_J93) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
# should look at resp failure
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_resp_failure_J96) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_resp_failure_J96) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_pneumothorax_J93) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_pneumothorax_J93) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_malnutrition) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_malnutrition) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_other_nutritional_deficiencies_E50_E64) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_other_nutritional_deficiencies_E50_E64) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_typeI_diabetes_E10) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_typeI_diabetes_E10) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_typeII_diabetes_E11) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_typeII_diabetes_E11) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_obesity_overweight_E66) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_obesity_overweight_E66) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_liver_disease_K70_K77) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_liver_disease_K70_K77) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
# this crazy to look at
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_other_GI_notLiver_K_excludesK70K77) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_other_GI_notLiver_K_excludesK70K77) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_essential_hypertension_I10) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_essential_hypertension_I10) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_hypertensive_heart_disease_I11) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_hypertensive_heart_disease_I11) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_chronic_kidney_disease_N18) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_chronic_kidney_disease_N18) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_nicotine_dependence_F17) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_nicotine_dependence_F17) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_heart_failure_I50) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_heart_failure_I50) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_ischemic_heart_disease_I20_I25) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_ischemic_heart_disease_I20_I25) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

```{r}
#patients with cf 
sub_df %>%
  group_by(COVIDseverity, comorb_lung_transplant_Z942) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()

# patients without cf
nonsub_df %>%
  group_by(COVIDseverity, comorb_lung_transplant_Z942) %>%
  summarize(N = n()) %>%
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>%
  knitr::kable()
```

=======
>>>>>>> 089534ca4cc664aa7085839caa93e7a6b7798931
## Part 4

```{r}
con_tab <- d %>% 
  select(COVIDseverity, has_cf, age_at_encounter)
```


```{r}
#contingency table, compares non-cf covid severity with race
non_cf %>% 
  group_by(COVIDseverity, encountertype) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(N / sum(N), 3)) %>% 
  knitr::kable()
```

```{r}
#contingency table, compares cf covid severity with encountertype
cf %>% 
  group_by(COVIDseverity, encountertype) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(N / sum(N), 3)) %>% 
  knitr::kable()
```

```{r}
#contingency table, compares cf covid severity with encountertype
non_cf %>% 
  group_by(COVIDseverity, encountertype) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(N / sum(N), 3)) %>% 
  knitr::kable()
```

```{r}
#contingency table, compares covid severity between cf and non-cf
con_tab %>% 
  group_by(has_cf, COVIDseverity) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round((N / sum(N)), 3)) %>% 
  knitr::kable()
```

```{r}
#contingency table, compares covid severity between cf and non-cf (shows percentage)
con_tab %>% 
#contingency table, compares covid severity between cf and non-cf
ratetable <- con_tab %>% 
  group_by(has_cf, COVIDseverity) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
ratetable <- con_tab %>% 
  group_by(has_cf, COVIDseverity) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round((N / sum(N)), 3))
```


```{r}
#bar plot, covid severity for cf patients
ggplot(ratetable, aes(x = has_cf,
                      y = Ratio,
                      fill = COVIDseverity), position = "dodge") +
    labs(x = "Patients Without CF", 
       y = "Rate", 
       title = "COVID Severity Among Children With/Without Cystic Fibrosis") +
  geom_bar(stat = "identity")
```

```{r}
library(ggplot2)
theme_set(theme_classic())

ratetable <-
```


```{r}
#bar plot, covid severity for non-cf patients
ggplot(non_cf) +
  geom_bar(mapping = aes(x=has_cf, fill = COVIDseverity), 
           position = 'dodge') +
  scale_y_continuous(labels = comma) +
  labs(x = "Patients Without CF", 
       y = "Count", 
       title = "COVID Severity Among Children Without Cystic Fibrosis")
```

```{r}
ggplot(con_tab) +
  geom_bar(mapping = aes(x=has_cf, fill = COVIDseverity), 
           position = 'dodge') +
  coord_cartesian(ylim = c(0, 60000)) +
  scale_y_continuous(labels = comma)
```

Logistic Regression
---------------------
```{r}
library(aod)
library(ggplot2)
library(mlbench)
library(MASS)
library(pROC)

```

```{r}
covsev01 <- d
```

```{r}
covsev01 <-
  d["COVIDseverity"][d["COVIDseverity"] == 2] <- 1
```

```{r}
covid_data2 <- d %>% 
  select(COVIDseverity, has_cf, age_at_encounter)
```

```{r}
covid_data2["COVIDseverity"][covid_data2["COVIDseverity"] == 2] <- 1
```

```{r}
model <- glm(COVIDseverity ~., family = binomial(link='logit'),
             data = covid_data2)
summary(model)
```

```{r}
confint(model)
```

```{r}
confint.default(model)
```

```{r}
wald.test(b = coef(model), Sigma = vcov(model), Terms = 1:2)
```

```{r}
exp(coef(model))
```

```{r}
exp(cbind(OR = coef(model), confint(model)))
```

```{r}
model_2 <- stepAIC(model)
```


```{r}
hist(model_2$fitted.values,main = " Histogram ",xlab = "Probability of 'pos' diabetes", col = 'light green')
```




```{r}
ggplot(covid_data2, aes(x = has_cf, y = COVIDseverity)) + geom_ribbon(aes(ymin = LL,
    ymax = UL, fill = age_at_encounter), alpha = 0.2) + geom_line(aes(colour = rank),
    size = 1)
```


## Part 5
```{r}
non_cf <- d %>%
  filter(has_cf == 0)
```

```{r}
cf <- d %>%
  filter(has_cf == 1)
```

```{r}
non_cf %>% 
  ggplot(aes(x=COVIDseverity)) +
  geom_bar(fill = "magenta")
```
```{r}
  d %>% 
  select(has_cf, COVIDseverity) %>% 
  table()
```
```{r}
non_cf %>% 
  select(age_at_encounter, ethnicity) %>% 
  table()
```
```{r}
non_cf %>%
 group_by(COVIDseverity) %>%
 count()
```
```{r}
non_cf %>% 
  select(COVIDseverity, ethnicity) %>% 
  table()
```
```{r}
non_cf %>% 
  select(COVIDseverity, zip_code) %>% 
  table()
```
```{r}
non_cf %>% 
  select(COVIDseverity, race) %>% 
  table()
```
```{r}
non_cf %>% 
  select(COVIDseverity, encountertype) %>% 
  table()
```
```{r}
non_cf %>% 
  ggplot(aes(x=COVIDseverity,
             y=age_at_encounter, fill = COVIDseverity)) +
  geom_boxplot() +
  labs(x = "COVID Severity", 
       y = "Age at Encounter", 
       title = "COVID Severity among children without Cystic Fibrosis") +
  scale_fill_manual(values =c("#E69F00","#009E73","#56B4E9"))
```
```{r}
non_cf %>% 
  ggplot(aes(x=COVIDseverity, fill = race)) +
  geom_bar()
```
```{r}
d %>% 
  ggplot(aes(x=COVIDseverity,
             y=age_at_encounter, 
             fill = has_cf)) +
  geom_boxplot() +
  labs(x = "COVID Severity", y = "Age at Encounter") +
  ggtitle("COVID Severity Among Children based on Age") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_manual(values =c("#E69F00","#009E73"), 
                    labels = c("Non CF","CF"), 
                    name = "Cystic Fibrosis Diagnosis") +
  scale_x_discrete(breaks=c("0", "1","2"),
                      labels=c("AsymptomaticMild", "Moderate","Severe"))
```
```{r}
non_cf %>% 
  ggplot(aes(x=race, fill = COVIDseverity)) +
  geom_bar(position = "dodge") + 
  coord_flip() +
  labs(x = "Race", y = "Count") +
  ggtitle("COVID Severity Among Children based on Race", subtitle = "Non Cystic Fibrosis") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle=element_text(hjust=0.5)) +
  scale_fill_manual(values =c("#E69F00","#009E73","#56B4E9"), 
                    labels = c("Asymptomatic","Moderate","Severe"),
                    name = "COVID Severity") +
  scale_y_continuous(labels = comma)
```
```{r}
cf %>% 
  ggplot(aes(x=race, fill = COVIDseverity)) +
  geom_bar(position = "dodge") + 
  coord_flip() +
  labs(x = "Race", y = "Count") +
  ggtitle("COVID Severity Among Children based on Race",
          subtitle = "Cystic Fibrosis") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle=element_text(hjust=0.5)) +
  scale_fill_manual(values =c("#E69F00","#009E73","#56B4E9"), 
                    labels = c("Asymptomatic","Moderate","Severe"),
                    name = "COVID Severity") +
  scale_y_continuous(labels = comma)
```

```{r}
non_cf %>% 
  ggplot(aes(x=ethnicity, fill = COVIDseverity)) +
  geom_bar(position = "dodge") +
  labs(x = "Ethnicity", y = "Count") +
  ggtitle("COVID Severity Among Children based on Ethnicity",
          subtitle = "Non Cystic Fibrosis") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle=element_text(hjust=0.5)) +
  scale_fill_manual(values =c("#E69F00","#009E73","#56B4E9"), 
                    labels = c("Asymptomatic","Moderate","Severe"),
                    name = "COVID Severity") +
  scale_y_continuous(labels = comma) 
```
```{r}
cf %>% 
  ggplot(aes(x=ethnicity, fill = COVIDseverity)) +
  geom_bar(position = "dodge") +
  labs(x = "Ethnicity", y = "Count") +
  ggtitle("COVID Severity Among Children based on Ethnicity",
          subtitle = "Cystic Fibrosis") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle=element_text(hjust=0.5)) +
  scale_fill_manual(values =c("#E69F00","#009E73","#56B4E9"), 
                    labels = c("Asymptomatic","Moderate","Severe"),
                    name = "COVID Severity") +
  scale_y_continuous(labels = comma)
```

```{r}
d %>% 
  ggplot(aes(x=encountertype, 
             fill = has_cf)) +
  geom_bar(position = "dodge") +
  labs(x = "Type of Encounter", y = "Count") +
  ggtitle("COVID Severity Among Children based on Type of Encounter") +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_fill_manual(values =c("#E69F00","#009E73"), 
                    labels = c("Non CF","CF"), 
                    name = "Cystic Fibrosis Diagnosis") +
  coord_flip() +
  scale_y_continuous(labels = comma)
```
```{r}
table(d$ethnicity,d$has_cf)
```
```{r}
table(d$race,d$COVIDseverity)
```

X AXIS 3 CATEGORIES Y AXIS 2 CATEGORIES
```{r}
table(d$ethnicity,d$encountertype)
```

```{r}
# Medication 1
table(d$COVIDseverity,d$INOTROPES)
```
```{r}
# cf
cf_intropes <- table(cf$COVIDseverity,cf$INOTROPES)
mosaicplot(cf_intropes)
```
```{r}
fisher.test(cf_intropes)
chisq.test(cf_intropes)
```
```{r}
# non cf
non_cf_intropes <- table(non_cf$COVIDseverity,non_cf$INOTROPES)
mosaicplot(non_cf_intropes)
```

```{r}
# contingency table comparing COVID severity and medication INOTROPES
cf %>% 
  group_by(COVIDseverity, INOTROPES) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
non_cf %>% 
  group_by(COVIDseverity, INOTROPES) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```

```{r}
table(d$COVIDseverity,d$REMDESIVIR)
```
```{r}
cf %>% 
  group_by(COVIDseverity, REMDESIVIR) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
non_cf %>% 
  group_by(COVIDseverity, REMDESIVIR) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
table(d$COVIDseverity,d$CCP)
```
```{r}
cf %>% 
  group_by(COVIDseverity, CCP) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
non_cf %>% 
  group_by(COVIDseverity, CCP) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
table(d$COVIDseverity,d$DEXAMETHASONE)
```
```{r}
cf %>% 
  group_by(COVIDseverity, DEXAMETHASONE) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
non_cf %>% 
  group_by(COVIDseverity, DEXAMETHASONE) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
table(d$COVIDseverity,d$ENOXAPARIN)
```
```{r}
cf %>% 
  group_by(COVIDseverity, ENOXAPARIN) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
non_cf %>% 
  group_by(COVIDseverity, ENOXAPARIN) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
table(d$COVIDseverity,d$HEPARIN)
```
```{r}
cf %>% 
  group_by(COVIDseverity, HEPARIN) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
non_cf %>% 
  group_by(COVIDseverity, HEPARIN) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
table(d$COVIDseverity,d$IVIG)
```
```{r}
cf %>% 
  group_by(COVIDseverity, IVIG) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
non_cf %>% 
  group_by(COVIDseverity, IVIG) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
table(d$COVIDseverity,d$METHYLPREDNISOLONE)
```
```{r}
cf %>% 
  group_by(COVIDseverity, METHYLPREDNISOLONE) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
non_cf %>% 
  group_by(COVIDseverity, METHYLPREDNISOLONE) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
table(d$COVIDseverity,d$RITUXIMAB)
```
```{r}
cf %>% 
  group_by(COVIDseverity, RITUXIMAB) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
non_cf %>% 
  group_by(COVIDseverity, RITUXIMAB) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
table(d$COVIDseverity,d$TOCILIZUMAB)
```
```{r}
cf %>% 
  group_by(COVIDseverity, TOCILIZUMAB) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
non_cf %>% 
  group_by(COVIDseverity, TOCILIZUMAB) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
table(d$COVIDseverity,d$ASPIRIN)
```
```{r}
cf %>% 
  group_by(COVIDseverity, ASPIRIN) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
non_cf %>% 
  group_by(COVIDseverity, ASPIRIN) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
table(d$COVIDseverity,d$LOPINAVIR_OR_RITONAVIR)
```
```{r}
cf %>% 
  group_by(COVIDseverity, LOPINAVIR_OR_RITONAVIR) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```
```{r}
non_cf %>% 
  group_by(COVIDseverity, LOPINAVIR_OR_RITONAVIR) %>% 
  summarize(N = n()) %>% 
  mutate(Ratio = round(100*(N / sum(N)), 1), "%") %>% 
  knitr::kable()
```












